### 2012-07-09 inst/examples/monitoring_impact_warning.Rmd
example of the impact that magagement has on ability to detect warning signals

 * Author [Carl Boettiger](http://carlboettiger.info), <cboettig@gmail.com>
 * License: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
 * Description:  Implements a numerical version of the SDP described in `r knitcitations::citet("10.1016/0095-0696(79)90014-7")`, in which the dynamics are slowly moving towards a tipping point.

@@

``` {r }
pars <- c(r = .75, k = 10, a=1, H=1, Q = 3)
K <- 8 # approx
````

Ask R to show us how this function is defined, and plot the transition point using these parameters.

``` {r showMay, fig.keep="last"}
May

curve(.75*(1-x/10), 0, 10)
curve(1*x^2/(x^3+1), 0, 10, add=T, col="blue")
curve(1.9*x^2/(x^3+1), 0, 10, add=T, col="red")
````

@@

Now we'll simulate 100 replicates of this stochastic process, but with the system dynamics gradually changing over time.

``` {r simulationfunction}
Dt <- 1

ForwardSimulate <- 
  function(f, pars, x_grid, h_grid, x0, D, z_g,
         z_m=function(x) 1, z_i = function(x) 1, 
         profit=NULL){

  OptTime <- dim(D)[2]    # Stopping time
  x_h <- numeric(OptTime) # population dynamics with harvest
  h <- numeric(OptTime) # optimal havest level
  x_h[1] <- x0  # initial values
  s <- x_h # also track escapement
  x <- x_h # What would happen with no havest
  p <- numeric(OptTime)

  ## Simulate through time ##
  for(t in 1:(OptTime-1)){

    # Move towards collapse
    pars[3] <- pars[3] + Dt/OptTime

    # Assess stock, with potential measurement error
    m_t <- x_h[t] * z_m()
    # Current state (is closest to which grid posititon) 
    St <- which.min(abs(x_grid - m_t)) 
    q_t <- h_grid[D[St, (t + 1) ]] 
    # Implement harvest/(effort) based on quota with noise 
    h[t] <- q_t * z_i()
    # Noise in growth 
    z <- z_g() 
    # population grows
    x_h[t+1] <- z * f(x_h[t], h[t], pars) # with havest
    s[t+1]   <- x_h[t] - q_t # anticipated escapement
    x[t+1]   <- z * f(x[t], 0, pars) # havest-free dynamics
    p[t] <- profit(x_h[t], h[t])
  }
  data.frame(time = 1:OptTime, fishstock = x_h, harvest = h,
             unharvested = x, escapement = s, profit = p) 
}

````

The bifurcation parameter will increase from `r pars[3]` to `r pars[3] + Dt*(OptTime-1)/OptTime`.

@@

The forward simulation algorithm needs an initial condition `x0` which we set equal to the carrying capacity, as well as our population dynamics `f`, parameters `pars`, grids, and noise coefficients.  Recall in the Reed case only `z_g`, growth, is stochastic.

@@

Let's begin by looking at the dynamics of a single replicate. The line shows Reed's S, the level above which the stock should be harvested (where catch should be the difference between stock and S).  To confirm that this policy is being followed, note that harvesting only occurs when the stock is above this line, and harvest is proportional to the amount by which it is above.  Change the replicate `reps==` to see the results from a different replicate.

@@

## Calculate warning signals

@@

# References

``` {r biblio, echo=FALSE, results="asis"}
knitcitations::bibliography("html")
````

>>>>

 * Author [Carl Boettiger](http://carlboettiger.info), <cboettig@gmail.com>
 * License: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
 * Description:

@@

``` {r }
pars <- c(r = .75, k = 10, a=1.7, H=1, Q = 3)
K <- 8 # approx
````

Note that this bifurcates when a increases to 2.

@@

@@

@@

@@

## Calculate warning signals on managed system

@@

## Unmanaged warning signals
Calculate warning signals on the same set of shocks in unfished data

``` {r earlywarning2}
acor_tau <- dt[unharvested > 1.5, 
               warningtrend(data.frame(time=time, value=unharvested),
                            window_autocorr),
               by=reps]

var_tau <- dt[unharvested > 1.5, 
              warningtrend(data.frame(time=time, value=unharvested),
                          window_var),
              by=reps]
````

``` {r modelbased2}
m <- dt[unharvested > 1.5, 
        stability_model(data.frame(time=time, value=unharvested),
                          "LSN")$pars["m"],
        by=reps]
````

``` {r summaryplot2}
signals <- melt(data.frame(var=var_tau$V1, acor=acor_tau$V1, m=m$V1))
ggplot(signals) + geom_histogram(aes(value)) + facet_wrap(~variable, scales="free")
````
